---
title: "Titanic Survival Analysis using Decision Tree"
author: "Rishabh Lavangad"
date: "23 November 2018"
output:
  pdf_document: default
  html_document: default
---

1. Import the data set in R, create a variable called titanic & store the data in it.
```{r}
titanic <- read.csv("Titanic.csv")
head(titanic)
```

2. Print the structure of dataset. 
```{r}
str(titanic)
```

3. Print last 6 rows of dataset.
```{r}
tail(titanic,6)
```

4.Create a barplot using ggplot with class on x-axis, frequency on y -axis & sex as the fill, write your observation. 

```{r}
library(ggplot2)
ggplot(titanic, aes(x = Pclass, fill = Sex)) + geom_bar()
```

OBSERVATION :-
As compared to first two Pclass's,the third Pclass has more number of male
and female count.


5. Use ggplot() to estimate your chances of servival from the distibution of sexes within the classes of the ship; Hint: use fact grids.
```{r}
ggplot(titanic, aes(x = Pclass, fill = Sex)) + geom_bar(position = "dodge") + facet_grid(~Survived)
```

6. Use ggplot() to estimate your chances of survival based on your age from the distribution of sexes within the class of the ship; Hint: Use the above plot & overlay age using jitter plot .
```{r}
posn.j <- position_jitter(0.5, 0)
ggplot(titanic,aes(x=factor(Pclass),y=Age,col=factor(Sex)))+
        geom_jitter(size=3,alpha=0.5,position=posn.j)+
        facet_grid(". ~ Survived")
```

7. Import the training dataset using the following link and print the structure to console "http://50.amazonaws.com/assets.datacamp.com/course/kaggle/train.csv"
```{r}
train_url <- "http://s3.amazonaws.com/assets.datacamp.com/course/Kaggle/train.csv"
train <- read.csv(train_url)
str(train)
```

8. Import the testing dataset using the following link & print the structure to console "http://50.amazonaws.com/assets.datacamp.com/course/kaggle/train.csv"
```{r}
test_url <- "http://s3.amazonaws.com/assets.datacamp.com/course/Kaggle/test.csv"
test <- read.csv(test_url)
str(test)
```

9. Print the survival rates in absolute numbers for train.
```{r}
table(train$Survived)
```

10. Print th survival rates in proportion for train.
```{r}
prop.table(table(train$Survived))
```

11. Print a two way comparision table for sex and survived
```{r}
table(train$Sex, train$Survived)
```

12. Print a two way comparision for sex and survived, show the proportions row wise i.e for Male & Female
```{r}
prop.table(table(train$Sex, train$Survived),margin =1)
```

13. Create a column called child using the condition age < 18 
```{r}
library(dplyr)
train$Child <- NA
train$Child[train$Age < 18] <- "1"
train$Child[train$Age >= 18] <- "0"
head(train)
```

14. Print a two way comparison table to show the proportion of this child variable with respect to survival
```{r}
prop.table(table(train$Child,train$Survived))
```

15. Using rpart build a decision tree on the training dataset 
Dependent Vaar : Survived 
Independent var : Pclass,sex,age,parch,fare,embarked

```{r}
library(caret)
train$Survived=as.factor(train$Survived)
tree <- train(Survived  ~ Pclass + Sex + Age + SibSp + Parch + Fare + Embarked, data = train, method = "rpart", na.action = na.exclude)
plot(tree$finalModel,uniform = T,main="decison tree")
text(tree$finalModel,use.n = T,all=T,cex=.8)
```


16. Plot your decision tee using fancyrpart plots
```{r}
library(rattle)
fancyRpartPlot(tree$finalModel)
```

17. Make predictions on the test set
```{r}
prediction <- predict(tree, test,type = "prob")
head(prediction)
```